Solving the maximum independant set problem by generic programming einsum networks

### 2021-06-11 paper/paper.tex
update

\iclrfinalcopy
\begin{document}
\title{Solving the maximum independant set problem with einsum networks}

@@

\begin{abstract}
	Solving the maximum independent set size problem by mapping the problem to an einsum network. 
    We obtain the maximum independent set size,
    the independence polynomial and the optimal configuration by using different tensor elements to a commutative semiring.
\end{abstract}

\section{Three approaches to computing independence polynomial}
The independence polynomial~\cite{Ferrin2014, Harvey2017} of a graph $G$ is defined as
\begin{equation}
I(G, x) = \sum_{k=1}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in $G$, and $\alpha(G)$ is the maximum independent set size.
\blue{here, I preassumed a user knows tensor network, add more details later.}
If we map a graph into an einsum network, as shown in \Fig{fig:einsummapping}, by placing a rank one tensor of size $2$ at each vertex
\begin{equation}
    W(x)_{s_i} = \left(\begin{matrix}
        1 \\
        x
    \end{matrix}\right)_{s_i},
\end{equation}
and a rank two tensor of size $2 \times 2$ at each edge
\begin{equation}
    B_{s_i s_j} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right)_{s_is_j},
\end{equation}
the contraction of this einsum network generates the independence polynomial %, we denote this contraction as \texttt{a,b,c,d,ab,ad,bc,bd,cd,ce->}.
\begin{equation}
    I(G, x) = \sum\limits_{s_1, s_2, \ldots, s_n = 0}^{1} \prod\limits_{i=1}^n W(x)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j}.
\end{equation}
Here, the einsum enumerates all possible configurations and accumulates the product of tensor elements to the output.
The product over vertex tensors provides a factor $x^k$, where $k=\sum_i s_i$ is the set size.
While the product over edge tensors provides a factor $0$ if the configuration is not an independent set, otherwise $1$.
\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.4\textwidth}
        \centering
        \centerline{\includegraphics[width=0.8\columnwidth,trim={0 0cm 0 0},clip]{../notebooks/fig1.pdf}}
        \caption{}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.4\textwidth}
        \centering
    \centerline{\includegraphics[width=0.8\columnwidth,trim={0 0cm 0 0},clip]{../notebooks/fig2.pdf}}
        \caption{}
    \end{subfigure}
    \caption{Mapping a graph to an einsum network.}\label{fig:einsummapping}
\end{figure*}
By mapping the problem to einsum, one can use some existing algorithms to decrease the complexity of the problem.
Especially those techiniques developed in recent quantum circuit simulations for finding the optimal exactly contraction order of a tensor network ~\cite{Gray2021,Pan2021}.
With these algorithms, one can easily find a good contraction order that reduces the complexity to a quantity related to the minimum treewidth of the graph.~\cite{Markov2008}
Such mapping also makes it possible to utilize faster BLAS functions.
Before contracting the tensor network and evaluating this polynomial numerically, let us first give up thinking $0$s and $1$s in tensors $W(x)$ and $B$ as regular computer numbers such as integers and floating point numbers.
Instead, we treat them as the additive identity and multiplicative identity of a commutative semiring.
A ring without additive inverse is called a semiring, and commutative semiring is a semiring that multiplication commutative.
To define a commutative semiring with addition algebra $\oplus$ and multiplication algebra $\odot$ on a set $R$, the following relation must hold for arbituary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b + a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} * a = \mymathbb{0}
\end{align*}

\subsection{Symbolic computing}
To evaluate this polynomial, one can use the symbolic computing directly, that is, storing a polynomial of $x$ as a vector of factors.
We define the algebra between the polynomials $a$ of order $k_a$ and $b$ of order $k_b$ as
\begin{align}
    \begin{split}
    a \oplus b &= \{a_0 + b_0, a_1 + b_1, \ldots, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}\},\\
    a \odot b &= \{a_0 + b_0, a_1b_0 + a_0b_1, \ldots, a_{k_a} b_{k_b}\},\\
    \mymathbb{0} &= \{\},\\
    \mymathbb{1} &= \{1\}.
    \end{split}
\end{align}
The variable $x$ now is represented as a vector $\{0, 1\}$.
Then we insert them to the original tensor network contraction algorithm, the final result is the exact representation of the independence polynomial.
In the program, the multiplication can be evaluated efficiently with the convolution theorem.
However, this method suffers from a space overhead that propotional to the maximum independant set size because each polynomial requires a vector of such size to store.

\subsection{Polynomial fitting and Fourier transformation}
Let $m=\alpha(G)$ be the maximum independent set size and $X$ be a set of $m+1$ random real numbers, e.g. $\{0, 1, 2, \ldots, m\}$.
We compute the einsum contraction for each $x \in X$ and obtain the following relations
\begin{align}
    \begin{split}
a_0 + a_1 x_1 + a_1 x_1^2 + \ldots + a_m x_1^m &= y_0\\
a_0 + a_1 x_2 + a_2 x_2^2 + \ldots + a_m x_2^m &= y_1\\
\ldots&\\
a_0 + a_1 x_m + a_2 x_m^2 + \ldots + a_m x_m^m& = y_m
    \end{split}
\end{align}
The polynomial fitting between $X$ and $Y = \{y_0, y_1, \ldots, y_m\}$ gives us the factors.
The polynomial fitting is esentially about solving the following linear equation
\begin{align}
\left(\begin{matrix}
1 & x_1 & x_1^2 & \ldots & x_1^m \\
1 & x_2 & x_2^2 & \ldots & x_2^m \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & x_m & x_m^2 & \ldots & x_m^m
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_m
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_m
\end{matrix}\right).
\end{align}

However, the result is very unreliable in practise due to the non-negligible round off error of floating point numbers.
Instead of choosing $X$ as a set of random real numbers, we let $x_i = r\omega^i$ and the above linear equation becomes
\begin{align}
\left(\begin{matrix}
1 & r\omega & r^2\omega^2 & \ldots & r^m\omega^m \\
1 & r\omega^2 & r^2\omega^4 & \ldots & r^m\omega^{2m} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & r\omega^m & r^2\omega^{2m} & \ldots & r^m\omega^{m^2}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_m
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_m
\end{matrix}\right).
\end{align}

When $r=1$, the left side is a DFT matrix. We can obtain the factors using the relation $\vec a = {\rm FFT^{-1}}(\omega) \cdot \vec y$.
In the special case that $\omega = e^{-2\pi i/(m+1)}$, it is directly solvable with inverse fast fourier transformation algorithm in package \texttt{FFTW}.

\begin{equation}
{\rm FFT}(\omega) \cdot \vec a_r = \vec y
\end{equation}
where $(\vec a_r)_k = a_k r ^k$, by choosing diferent $r$, we can obtain better precision in low independant set size region  ($\omega<1$) and high independant set size region ($\omega>1$).

\subsection{Finite field algebra}
It is possible to compute the independence polynomials exactly using basic integer types only,
even when the degeneracy is bigger than that can be represented by 64 bit integers.
What we need is computing with the finite field algebra $GF(p)$

\begin{align}
\begin{split}
    x\pmod p ~\oplus~ y\pmod p &= x+y\pmod p\\
    x\pmod p ~\odot~ y\pmod p &= xy\pmod p\\
    \mymathbb{0} &= 0\\
    \mymathbb{1} &= 1
\end{split}
\end{align}

It a field because the multiplicative inverse can defined on non-zero values, which can be computed with the extended Euclidean algorithm.
In a finite field algebra, we have the following observations
\begin{enumerate}
    \item One can still use Gaussian elimination~\cite{Golub2013} to solve the linear equations in a finite field, since the multiplicative inverse is properly defined in a finite field,
    \item Let remainders of an integer $x$ over a set of coprime integers $\{p_1, p_2, \ldots, p_n\}$ being $\{x \pmod {p_1}, x\pmod {p_2}, \ldots, x \pmod {p_n}\}$,
    then $\{x \pmod {p_1 \times p_2 \times \ldots \times p_n}\}$ can be obtained using the chinese remainder theorem.
\end{enumerate}
With these observations, we use Algorithm~\ref{alg:finitefield} to compute degeneracies.
In the algorithm, except the computation of chinese remainder theorem, all computations are done with integers with fixed width.

\begin{algorithm}[!ht]
    \small
    \SetAlgoNoLine
    \LinesNumbered
    Let $P = 1$, $X = {0,1,2,\ldots,m}$ and $\hat X_{ij} = X_i^j$, where $i,j = \{0, 1, \ldots m\}$\;
    \While{true}{
        compute the largest prime 64 bit integer $p$ that $\gcd(p, P) = 1$\;
        compute the tensor network contraction on $GF(p)$ and obtain $Y\pmod p = \{y_0 \pmod p, y_1 \pmod p, \ldots , y_m \pmod p\} $\;
        $A \pmod p = \{a_0 \pmod p, a_1 \pmod p, \ldots, a_m \pmod p\} = {\rm gaussian\_elimination}(\hat X, Y \pmod p) $\;
        $A \pmod {P\times p} = {\rm chinese\_remainder}(A \pmod P, A \pmod p)$\;
        \If{$A \pmod P = A \pmod {P \times p}$}{
            \Return $A \pmod P$\;
        }
        $P = P \times p$\;
    }\caption{Compute independence polynomial exactly without integer overflow}\label{alg:finitefield}
\end{algorithm}

\section{Computing maximum independent set size and its corresponding degeneracy and configurations}
Obtaining the maximum independent set size and its degeneracy is computational more efficient. Let $x=\infty$, then the independence polynomial becomes
\begin{equation}
I(G, \infty) = a_k \infty^{\alpha(G)},
\end{equation}
where the lower orders terms disappear automatically. We can define a new algebra as
\begin{align}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}\\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 0\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}
\end{split}
\end{align}
In the program, we only store the power $x$ and the corresponding factor $a_x$ that initialized to $1$.
This algebra is consistent with the one we derived in ~\cite{Liu2021} that uses the tropical tensor network solving spin glass ground states.
Here if we are only interested in obtaining $\alpha(G)$ and omit the factors, the algebra of $x$ corresponds to the tropical algebra~\cite{Maclagan2015,Moore2011}.

One may want to obtain all ground state configurations,
it can be achieved replacing the factors $a_x$ with a set of bit strings $s_x$ initialized to a vertex index $i$ dependent onehot vector $s_i = \{\boldsymbol{\delta}_{ij=1,\ldots,n}\}$.
\iffalse
\begin{align}
\begin{split}
    s_x \oplus s_y &= s_x \cup s_y\\
    s_x \odot s_y &= \{\sigma \lor \tau | \sigma \in s_x, \tau \in s_y\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{\boldsymbol 0\}
\end{split}
\end{align}
\fi
\begin{align}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases}\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y}\\
    \mymathbb{0} &= \{\}\infty^{-\infty}\\
    \mymathbb{1} &= \{\boldsymbol 0\}\infty^{0}
\end{split}
\end{align}
The $\vee$ is the bit-wise \texttt{or}. One can easily check that this replacement does not change the fact that the algebra is a commutative semiring.
By slightly modifying the above algebra, it can also be used to obtain just a single configuration to save computational effort.
We leave this as an exercise for readers.

>>>>

\iclrfinalcopy
\begin{document}
\title{Solving the maximum independant set problem by generic programming einsum networks}

@@

\begin{abstract}
	Solving the maximum independent set size problem by mapping the graph to an einsum network. 
    By contracting the einsum network with generic element types, we show how to obtain the maximum independent set size,
    the independence polynomial and optimal configurations.
\end{abstract}

\section{Independence polynomial}
The independence polynomial~\cite{Ferrin2014, Harvey2017} of a graph $G$ is defined as
\begin{equation}
I(G, x) = \sum_{k=1}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in $G$, and $\alpha(G)$ is the maximum independent set size.
\blue{here, I preassumed a user knows tensor network, add more details later.}
Let us map the graph $G$ into an einsum network, as shown in \Fig{fig:einsummapping}, by placing a rank one tensor of size $2$ on vertex $i$
\begin{equation}
    W(x)_{s_i} = \left(\begin{matrix}
        1 \\
        x
    \end{matrix}\right)_{s_i},
\end{equation}
and a rank two tensor of size $2 \times 2$ on edge $(i,j)$
\begin{equation}
    B_{s_i s_j} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right)_{s_is_j},
\end{equation}
where a tensor index $s_i$ is a boolean variable that being 1 if vertex $i$ is in the independent set, 0 otherwise. The contraction of this einsum network gives the independence polynomial %, we denote this contraction as \texttt{a,b,c,d,ab,ad,bc,bd,cd,ce->}.
\begin{equation}
    I(G, x) = \sum\limits_{s_1, s_2, \ldots, s_n = 0}^{1} \prod\limits_{i=1}^n W(x)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j}.
\end{equation}
Here, the einsum runs over all possible vertex configurations and accumulates the product of tensor elements to the output.
The product over vertex tensors provides a factor $x^k$, where $k=\sum_i s_i$ is the vertex set size, 
while the product over edge tensors provides a factor $0$ for configurations not being an independent set.
\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.4\textwidth}
        \centering
        \centerline{\includegraphics[width=0.8\columnwidth,trim={0 0cm 0 0},clip]{../notebooks/fig1.pdf}}
        \caption{}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.4\textwidth}
        \centering
    \centerline{\includegraphics[width=0.8\columnwidth,trim={0 0cm 0 0},clip]{../notebooks/fig2.pdf}}
        \caption{}
    \end{subfigure}
    \caption{Mapping a graph to an einsum network.}\label{fig:einsummapping}
\end{figure*}
With this einsum network representation, one can take the advantage of recently developed techiniques in tensor network based quantum circuit simulations ~\cite{Gray2021,Pan2021},
where people evaluate a tensor network by pairwise contracting tensors in a heuristic order.
A good contraction order can reduce the time complexity significantly, at the cost of having a space overhead of $O(2^{tw(G)})$, where $tw(G)$ is the treewidth of $G$.~\cite{Markov2008}
The pairwise tensor contraction also makes it possible to utilize fast basic linear algebra subprograms (BLAS) functions for certain tensor element types.

Before contracting the einsum network and evaluating the independence polynomial numerically, let us first give up thinking $0$s and $1$s in tensors $W(x)$ and $B$ as regular computer numbers such as integers and floating point numbers.
Instead, we treat them as the additive identity and multiplicative identity of a commutative semiring.
A semiring is a ring without additive inverse, while a commutative semiring is a semiring that multiplication commutative.
To define a commutative semiring with addition algebra $\oplus$ and multiplication algebra $\odot$ on a set $R$, the following relation must hold for arbituary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b + a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} * a = \mymathbb{0}
\end{align*}

In the rest of this paper, we show how to obtain the independence polynomial, the maximum independent set size and optimal configurations of a general graph $G$ by designing tensor element types as commutative semirings,
i.e. making the einsum network programming generic~\cite{Stepanov2014}.

\subsection{The polynomial approach}
A straight forward approach to evaluate the independence polynomial is treating the tensor elements as polynomials, and evaluate the polynomial directly.
Let us create a polynomial type, and represent a polynomial $a_0 + a_1 x + \ldots + a_k x^k$ as a vector $(a_0, a_1, \ldots, a_k) \in R^k$, e.g. $x$ is represented as $(0, 1)$.
We define the algebra between the polynomials $a$ of order $k_a$ and $b$ of order $k_b$ as
\begin{align}
    \begin{split}
    a \oplus b &= (a_0 + b_0, a_1 + b_1, \ldots, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_0 + b_0, a_1b_0 + a_0b_1, \ldots, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (),\\
    \mymathbb{1} &= (1).
    \end{split}
\end{align}
By contracting the einsum network with polynomial type, the final result is the exact representation of the independence polynomial.
In the program, the multiplication can be evaluated efficiently with the convolution theorem.
The only problem of this method is it suffers from a space overhead that propotional to the maximum independant set size because each polynomial requires a vector of such size to store the factors.
In the following subsections, we managed to solve this problem.

\subsection{The fitting and Fourier transformation approaches}
Let $m=\alpha(G)$ be the maximum independent set size and $X$ be a set of $m+1$ random real numbers, e.g. $\{0, 1, 2, \ldots, m\}$.
We compute the einsum contraction for each $x_i \in X$ and obtain the following relations
\begin{align}
    \begin{split}
a_0 + a_1 x_1 + a_1 x_1^2 + \ldots + a_m x_1^m &= y_0\\
a_0 + a_1 x_2 + a_2 x_2^2 + \ldots + a_m x_2^m &= y_1\\
\ldots&\\
a_0 + a_1 x_m + a_2 x_m^2 + \ldots + a_m x_m^m& = y_m
    \end{split}
\end{align}
The polynomial fitting between $X$ and $Y = \{y_0, y_1, \ldots, y_m\}$ gives us the factors.
The polynomial fitting is esentially about solving the following linear equation
\begin{align}
\left(\begin{matrix}
1 & x_1 & x_1^2 & \ldots & x_1^m \\
1 & x_2 & x_2^2 & \ldots & x_2^m \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & x_m & x_m^2 & \ldots & x_m^m
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_m
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_m
\end{matrix}\right).
\end{align}

In practise, the fitting can suffer from the non-negligible round off errors of floating point operations and produce unreliable results.
This is because the factors of independence polynomial can be different in magnitude by many orders.
Instead of choosing $X$ as a set of random real numbers, we make it form a geometric sequence in the complex domain $x_j = r\omega^j$, where $r \in \mathbb{R}$ and $\omega = e^{-2\pi i/(m+1)}$. The above linear equation becomes
\begin{align}
\left(\begin{matrix}
1 & r\omega & r^2\omega^2 & \ldots & r^m\omega^m \\
1 & r\omega^2 & r^2\omega^4 & \ldots & r^m\omega^{2m} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & r\omega^m & r^2\omega^{2m} & \ldots & r^m\omega^{m^2}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_m
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_m
\end{matrix}\right).
\end{align}

Let us rearange the factors $r^j$ to $a_j$, the matrix on left side is exactly the a descrete fourier transformation (DFT) matrix.
Then we can obtain the factors using the inverse fourier transformation $\vec a_r = {\rm FFT^{-1}}(\omega) \cdot \vec y$, where $(\vec a_r)_j = a_j r ^j$.
By choosing diferent $r$, one can obtain better precision in low independant set size region  ($\omega<1$) and high independant set size region ($\omega>1$).

\subsection{The finite field algebra approach}
It is possible to compute the independence polynomials exactly using 64 bit integers types only,
even when the factors are larger than that can be represented by 64 bit integers.
We achieve this by designing a finite field algebra $GF(p)$

\begin{align}
\begin{split}
    x ~\oplus~ y &= x+y\pmod p,\\
    x ~\odot~ y &= xy\pmod p,\\
    \mymathbb{0} &= 0,\\
    \mymathbb{1} &= 1.
\end{split}
\end{align}

In a finite field algebra, we have the following observations
\begin{enumerate}
    \item One can still use Gaussian elimination~\cite{Golub2013} to solve a linear equation.
    This is because a field has the property that the multiplicative inverse exists for any non-zero value.
    The multiplicative inverse here can be computed with the extended Euclidean algorithm.
    \item Given the remainders of a larger integer $x$ over a set of coprime integers $\{p_1, p_2, \ldots, p_n\}$,
    $x \pmod {p_1 \times p_2 \times \ldots \times p_n}$ can be computed using the chinese remainder theorem.
    With this, one can infer big integers even though its bit width is larger than the register size.
\end{enumerate}
With these observations, we developed Algorithm~\ref{alg:finitefield} to compute independent polynomial exactly without introducing space overheads.
In the algorithm, except the computation of chinese remainder theorem, all computations are done with integers with fixed width $W$.

\begin{algorithm}[!ht]
    \small
    \SetAlgoNoLine
    \LinesNumbered
    Let $P = 1$, vector $X = (0,1,2,\ldots,m)$, matrix $\hat X_{ij} = X_i^j$, where $i,j = 0, 1, \ldots m$\;
    \While{true}{
        compute the largest prime $p$ that $\gcd(p, P) = 1 \land p \leq 2^W$\;
        compute the tensor network contraction on $GF(p)$ and obtain $Y = (y_0, y_1, \ldots , y_m) \pmod p $\;
        $A_p = (a_0, a_1, \ldots, a_m) \pmod p = {\rm gaussian\_elimination}(\hat X, Y \pmod p) $\;
        $A_{P\times p} = {\rm chinese\_remainder}(A_P, A_p)$\;
        \If{$A_P = A_{P \times p}$}{
            \Return $A_P$ \tcp*[l]{converged}
        }
        $P = P \times p$\;
    }\caption{Compute independence polynomial exactly without integer overflow}\label{alg:finitefield}
\end{algorithm}

\section{Computing maximum independent set size and its corresponding degeneracy and configurations}
Obtaining the maximum independent set size and its degeneracy can be computational more efficient. Let $x=\infty$, then the independence polynomial becomes
\begin{equation}
I(G, \infty) = a_k \infty^{\alpha(G)},
\end{equation}
where the lower orders terms disappear automatically. We can define a new algebra as
\begin{align}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}\\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 0\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}
\end{split}
\end{align}
In the program, we only store the power $x$ and the corresponding factor $a_x$ that initialized to $1$.
This algebra is consistent with the one we derived in ~\cite{Liu2021} that uses the tropical tensor network for solving spin glass ground states.
If one is only interested in obtaining $\alpha(G)$, he can drop the factor parts, then the algebra of $x$ becomes the max-plus tropical algebra~\cite{Maclagan2015,Moore2011}.

One may also want to obtain all ground state configurations, it can be achieved replacing the factors $a_x$ with a set of bit strings $s_x$.
We design a new element type that having algebra
\iffalse
\begin{align}
\begin{split}
    s_x \oplus s_y &= s_x \cup s_y\\
    s_x \odot s_y &= \{\sigma \lor \tau | \sigma \in s_x, \tau \in s_y\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{\boldsymbol 0\}
\end{split}
\end{align}
\fi
\begin{align}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{\boldsymbol 0\}\infty^{0},
\end{split}
\end{align}
where $\vee$ is the bit-wise \texttt{or}. One can easily check that this replacement does not change the fact that the algebra is a commutative semiring.
We first initialize the bit strings of the variable $x$ in the vertex tensor to a vertex index $i$ dependent onehot vector $x_i = \boldsymbol{e}_{i}$,
then we contract the tensor network. The resulting object will give us the set of all optimal configurations.
By slightly modifying the above algebra, it can also be used to obtain just a single configuration to save computational effort.
We leave this as an exercise for readers.

\subsection{bounding the enumeration space}
When we try to implement the above algebra for enumerating configurations, we find the space overhead is larger than than we have expected.
It stores more than nessesary intermedite configurations. To speed up the computation, we use $\alpha(G)$ that much easier to compute for bounding.
We first compute the value of $\alpha(G)$ with tropical numbers and cache all intermediate tensors.
Then we compute a boolean masks for each cached tensor, where we use a boolean true to represent a tensor element having contribution to the maximum independent set (i.e. with a nonzero gradient) and boolean false otherwise.
Finally, we perform masked matrix multiplication using the new element type with the above algebra for obtaining all configurations.
To compute the masks, we ``back propagate'' the masks step by step through contraction process using the cached intermediate tensors.
Consider a tropical matrix multiplication $C = A B$, we have the following inequality
\begin{align}
    A_{ij} \odot B_{jk} &\leq C_{ik}.
\end{align}

Moving $B_{ik}$ to the right hand side, we have
\begin{align}
    A_{ij} &\leq (\oplus_{k} (C_{ik}^{-1} \odot B_{jk}))^{-1}
\end{align}
where the tropical multiplicative inverse is defined as the additive inverse of the regular algebra. The equality holds if and only if element $A_{ij}$ contributions to $C$ (i.e. has nonzero gradient).
Let the mask for $C$ being $\overline C$, the backward rule for gradient masks reads
\begin{align}
\overline{A}_{ij} = \delta(A_{ij}, ((C^{\circ-1} \circ \overline C )B^T)_{ij}^{\circ -1}),
\end{align}
where ${}^{\circ -1}$ is the Hadamard inverse, $\circ$ is the Hadamard multiplication, boolean false is treated as tropical zero and boolean true is treated as tropical one.
This rule defined on matrix multiplication can be easily generalized to the einsum of two tensors by replacing the matrix multiplication between $C^{\circ-1} \circ \overline C$ and $B^T$ by an einsum.
